---
title: "Data Exploration Graphs"
authors: "Claire Morgan, Lindsey Stowell, Madison McDonald, Natalie Meyers"
output: html_document
---

## Load in Packages and Data

Load and install packages that are needed for creating graphs and visualizing data.

```{r}

#install.packages("tidyverse")
#install.packages("here")

library(tidyverse)
library(here)

```

# Temperature Data

Graphs that explore the daily and monthly temperature data sets.

## Data Type Manipulation

Import processed daily temperature information and ensure all dates are date types.

```{r}

# Import Data

daily_temperature_data <- read_csv(here("data","processed","daily_temp_data.csv"))

monthly_temperature_data <- read_csv(here("data", "processed", "monthly_temp_data.csv"))

```

Create year-month in monthly temperature data
```{r}

monthly_temperature_data <- monthly_temperature_data %>% 
  mutate(month_date = make_date(year, month))

```


## Exploratory Graphs

### Daily Mean Temperatures

Line graph visualizing how mean daily temperature changed throughout all dates.

```{r}

ggplot(
  data = daily_temperature_data,
  aes(
    x = startDateTime, 
    y = mean_temp,
    color = siteID) 
  ) +
  geom_line(size = 0.2) +
theme_bw()

```

### Monthly Mean Temperature

```{r}

ggplot(
  data = monthly_temperature_data,
  aes(
    x = month_date, 
    y = mean_temp,
    color = siteID) 
  ) +
  geom_line(size = 0.5) +
theme_bw()

```

### Monthly Temperature Range

```{r}

# Summarize and group monthly temperature data

monthly_temp_range <- monthly_temperature_data |>
  group_by(siteID, month) |>
  summarize(
    mean = mean(mean_temp, na.rm = TRUE),
    max = mean(max_temp, na.rm = TRUE),
    min = mean(min_temp, na.rm = TRUE)
  )

ggplot(
  monthly_temp_range,
  aes(
    x = month,
    y = mean
  )) +
geom_pointrange(
  aes(
    ymin = min,
    ymax = max
  )
) +
  facet_wrap(~siteID) +
  labs(y = "Range of mean monthly temperatures (C)") +
  scale_x_continuous(limits = c(1, 12),
                     breaks = seq(1, 12, 1)) +
  theme_bw()

```
------------------------------------------------------------------------
### Daily Sum of Pathogen Status

```{r}

pathogen_status_data <- read_csv(here("data","processed","pathogen_status_data.csv"))

```

Filter to keep only positive pathogen results

```{r}

positive_pathogen <- pathogen_status_data %>% 
  filter(testResult == "Positive")

#confusion on what "HardTick DNA Quality" in the "testPathogenName" column means and wondering if we can filter it out since it doesn't seem like that is a pathogen

# KF: Yes, I think you want to omit it; it is a quality control indicator column

```

Summarize with counts of positive pathogen status each month

``` {r}

# Tick Count Graph

daily_positive_pathogen <- positive_pathogen |>
  mutate(
    month = month(collectDate)
  ) |>
  group_by(siteID, month) |>
  summarize(
    positiveCount = n()
  )

ggplot(
  data = daily_positive_pathogen,
  aes(
    x = month, 
    y = positiveCount) 
  ) +
  geom_point() +
  facet_wrap(~siteID) +
theme_bw()

```

----------------------------------------------------------
## Make a Map of NEON Data Sites

```{r, methods figure}

## Load Packages
#install.packages("maps")
library(here)
library(tidyverse)
library(maps)

## Read in metadata for all NEON sites
sites_all <- read_csv(here("data", "raw", "NEON_Field_Site_Metadata_20210226_0.csv")) %>%
  select(domainID:colocated_site, latitude:longitude, state:country)

## Read in pathogen positive proportion data (monthly)
monthly_pathogens <- read_csv(here("data","processed","monthly_pathogen_count.csv"))

## Define the list of our project's sites
mySites <- monthly_pathogens %>%
  distinct(siteID) %>%
  pull(siteID)
  
## Subset "sites" object for our project's sites
sites <- sites_all %>%
  filter(siteID %in% mySites) %>%
  arrange(siteID)

## List state names of our sites
myStates <- as.character(sites$state)

## Add pathogen info per site
site_summary <- monthly_pathogens %>%
  group_by(siteID) %>%
  summarize(
    average = mean(positiveProportion)
  )

## Join summary-site and create categories of prevalence
sites_plot <- sites %>%
  left_join(site_summary, by = "siteID")

sites_plot <- sites_plot %>%
  mutate(category = case_when(
    average == 0.00 ~ "0 pathogen",
    average < 0.05 ~ "<5% pathogen",
    average >= 0.05 ~ "≥5% pathogen"
  ))

## get US map with state boundaries
us_states <- map_data("state")

```

```{r}

## Map our sites
ggplot() +
  # US state outlines
  geom_polygon(data = us_states,
               aes(x = long, y = lat, group = group),
               fill = NA,
               color = "black",
               linewidth = 0.3) +
  # Site points
  geom_point(data = sites_plot,
             aes(x = longitude, y = latitude,
                 color = category),
             size = 2.5) +
  # Assign Colors to Category
  scale_color_manual(values = c("royalblue","pink","maroon"),
                     labels = c("0 pathogen","<5% pathogen","≥5% pathogen")) +
  coord_fixed(1.3) +
  theme_void() +
  labs(color = "Pathogen Status")

```