---
title: "Data Processing"
output: html_document
authors: "Claire Morgan, Lindsey Stowell, Madison McDonald, Natalie Meyers"
---

# Import and Load Packages

Import and load the necessary packages.

```{r}

#install.packages("tidyverse")
#install.packages("here")
#install.packages("dplyr")
#install.packages("lubridate")


library(tidyverse)
library(here)
library(dplyr)
library(lubridate)

```

------------------------------------------------------------------------

# Relative Humidity Data

Manipulate the relative humidity data to only have data we need and then create new data sets for daily temperatures and monthly temperatures.

## Import Raw CSV

Isolate needed fields and save the updated data set to our `processed` folder. The relative humidity data set does not exist in our `raw` folder because the data set is too large to upload.

```{r}

# Access humidity data CSV and select useful columns
# WARNING: file not stored in GitHub because raw data file is too large :(

humidity_data <- read_csv(here("data","raw","relative_humidity_data.csv"))

temperature_data <- humidity_data %>% 
  select(siteID, startDateTime, endDateTime, tempRHMean)

temperature_data %>% 
  write_csv(here("data","processed","temperature_data.csv"))

```

## Process Data

We are processing the data in two steps because the "processed" data file was pre-loaded based off the large data set from humidity data that does not exist in the GitHub. To avoid re-downloading the data, we are using the "processed" data set that is not manipulated.

```{r}

temperature_data <- read_csv(here("data","processed","temperature_data.csv"))

# Isolate needed columns, assign date data type, and filter NAs out of data

temperature_data <- temperature_data |>
  select(
    siteID, 
    startDateTime, 
    tempRHMean) |>
  mutate(
    startDateTime = as_date(startDateTime),
    month = month(startDateTime),
    year = year(startDateTime)
  ) |>
  filter(
    !is.na(tempRHMean) & !is.na(startDateTime)
  )

```

## Create Daily Temperature Data

Create a table that includes mean, max, and minimum daily temperatures.

```{r}

# Create table with daily mean, max, min temperatures

daily_temp_data <- temperature_data |> 
  group_by(siteID, startDateTime) |>
  summarize(
    mean_temp = mean(tempRHMean, na.rm = TRUE),
    max_temp = max(tempRHMean, na.rm = TRUE),
    min_temp = min(tempRHMean, na.rm = TRUE)
  ) 

# Save new data set to processed folder

daily_temp_data |> write_csv(here("data","processed","daily_temp_data.csv"))

```

## Create Monthly Temperature Data

Create a table that includes mean, max, and minimum monthly temperatures.

```{r}

# Create table with monthly/year mean, max, min temperatures

monthly_temp_data <- temperature_data |>
  group_by(siteID, year, month) |>
  summarize(
    mean_temp = mean(tempRHMean, na.rm = TRUE),
    max_temp = max(tempRHMean, na.rm = TRUE),
    min_temp = min(tempRHMean, na.rm = TRUE)
  )

# Save data set to processed folder

monthly_temp_data |> write_csv(here("data","processed","monthly_temp_data.csv"))

```

------------------------------------------------------------------------

# Tick Abundance Data

Manipulate the tick abundance data to only have data we need and ensure there is no "NA" data.

## Process Data

```{r}

# Access tick abundance data CSV

abundance_data <- read_csv(here("data","raw","tick_abundance_data.csv"))

# Select only useful columns, change date to date data type, add month/year columns, and filter NAs

abundance_data <- abundance_data %>% 
  select(
    siteID, plotID, nlcdClass, elevation, collectDate, sampleID, adultCount, nymphCount, larvaCount) %>%
  mutate(
    collectDate = as_date(collectDate),
    month = month(collectDate),
    year = year(collectDate)
  ) %>%
  drop_na(
    sampleID , adultCount , nymphCount , larvaCount
  )

# Save processed data to processed data folder

abundance_data %>% 
  write_csv(here("data","processed","abundance_data.csv"))

```

## Create Monthly Abundance Data

Create a table that includes a count of ticks for each month by life stage (larva/nymph/adult)

```{r}

# Create table with monthly count of ticks

monthly_tick_abundance <- abundance_data |>
  group_by(siteID, year, month) |>
  summarize(
    larvaCount = sum(larvaCount),
    nympthCount = sum(nymphCount),
    adultCount = sum(adultCount)
  )

# Save data set to processed folder

monthly_tick_abundance |> write_csv(here("data","processed","monthly_abundance_data.csv"))

```

------------------------------------------------------------------------

# Tick Pathogen Status Data

Manipulate the tick pathogen status data set to only have data we need, add month/year columns, and ensure there is no "NA" data.

## Process Data

```{r}

# Select only useful columns, change date to date data type, and filter out NAs and test for hardtick DNA quality (not a pathogen)

pathogen_status_data <- read_csv(here("data","raw","tick_pathogen_status_data.csv"))

pathogen_status_data <- pathogen_status_data %>% 
  select(
    siteID, plotID, nlcdClass, elevation, collectDate, individualCount, testResult, testPathogenName, subsampleID) %>% 
  mutate(
    collectDate = as_date(collectDate),
    month = month(collectDate),
    year = year(collectDate)
    ) %>% 
  filter(
    !is.na("testResult") & !is.na("collectDate") & testPathogenName != "HardTick DNA Quality"
    )

pathogen_status_data %>% 
  write_csv(here("data","processed","pathogen_status_data.csv"))

```

## Create Monthly Positive Status Data

Create a table that includes a count of positive pathogen tests each month.
**Note:** Pathogen status is only tested on up to 130 nymphs.

```{r}

# Create table with monthly count of positive tests

monthly_pathogens <- pathogen_status_data |>
  group_by(siteID, year, month) |>
  filter(testResult == "Positive") |>
  summarize(
    pathogenCount = n()
  )

# Save data set to processed folder

monthly_pathogens |> write_csv(here("data","processed","monthly_pathogen_count.csv"))

```

## Create Daily Positive Status Data

Create a table that includes a count of positive pathogen tests each day.
**Note:** Pathogen status is only tested on up to 130 nymphs.

```{r}
# Open pathogen status data

pathogen_status_data <- read_csv(here("data","processed","pathogen_status_data.csv"))

# Create table with daily count of positive tests

daily_pathogens <- pathogen_status_data |>
  group_by(siteID, collectDate) |>
  filter(testResult == "Positive") |>
  summarize(
    pathogenCount = n()
  )

# Save data set to processed folder

daily_pathogens |> write_csv(here("data","processed","daily_pathogen_count.csv"))

```

------------------------------------------------------------------------

# Temperature vs Pathogen Table

### We are creating a table to show temperature and the number of positive pathogens at each temperature point.

```{r}

# Read abundance csv and daily temperature csv

pathogens <- read_csv(here("data","processed","daily_pathogen_count.csv"))

daily_temp <- read_csv(here("data","processed","daily_temp_data.csv"))

daily_temp <- daily_temp %>% 
  rename(collectDate = startDateTime)

# Common columns between daily_temp and pathogens data

names(pathogens) %>%
  intersect(names(daily_temp))

pathogen_vs_temp <- full_join(daily_temp, pathogens) 

pathogen_vs_temp <- pathogen_vs_temp %>% 
  drop_na(pathogenCount)

pathogen_vs_temp |> write.csv(here("data","processed","pathogen_vs_temp_data.csv"))

```