---
title: "Data Exploration Graphs"
date: "2026-03-03"
output: html_document
---

## Load in Packages and Data

Load and install packages that are needed for creating graphs and visualizing data.

```{r}

#install.packages(dplyr)
#install.packages(ggplot2)
#install.packages(here)

library(tidyverse)
library(here)

```

# Temperature Data

Graphs that explore the daily and monthly temperature data sets.

## Data Type Manipulation

Import processed daily temperature information and ensure all dates are date types.

```{r}

daily_temperature_data <- read_csv(here("data","processed","daily_temp_data.csv"))

monthly_temperature_data <- read_csv(here("data", "processed", "monthly_temp_data.csv"))

```

Create year-month in monthly temperature data
```{r}

monthly_temperature_data <- monthly_temperature_data %>% 
  mutate(month_date = make_date(year, month))

```


## Exploratory Graphs

### Daily Mean Temperatures

Line graph visualizing how mean daily temperature changed throughout all dates.

```{r}

ggplot(
  data = daily_temperature_data,
  aes(
    x = startDateTime, 
    y = mean_temp,
    color = siteID) 
  ) +
  geom_line(size = 0.2) +
theme_bw()

```

### Monthly Mean Temperature

```{r}

ggplot(
  data = monthly_temperature_data,
  aes(
    x = month_date, 
    y = mean_temp,
    color = siteID) 
  ) +
  geom_line(size = 0.5) +
theme_bw()

```

### Monthly Temperature Range

```{r}

# Summarize and group monthly temperature data

monthly_temp_range <- monthly_temperature_data |>
  group_by(siteID, month) |>
  summarize(
    mean = mean(mean_temp, na.rm = TRUE),
    max = mean(max_temp, na.rm = TRUE),
    min = mean(min_temp, na.rm = TRUE)
  )

ggplot(
  monthly_temp_range,
  aes(
    x = month,
    y = mean
  )) +
geom_pointrange(
  aes(
    ymin = min,
    ymax = max
  )
) +
  facet_wrap(~siteID)

```
------------------------------------------------------------------------
### Daily Sum of Pathogen Status

```{r}

pathogen_status_data <- read_csv(here("data","processed","pathogen_status_data.csv"))

```

Filter to keep only positive pathogen results

```{r}

positive_pathogen <- pathogen_status_data %>% 
  filter(testResult == "Positive")

#confusion on what "HardTick DNA Quality" in the "testPathogenName" column means and wondering if we can filter it out since it doesn't seem like that is a pathogen

```

Summarize with counts of positive pathogen status each day

``` {r}


daily_positive_pathogen <- positive_pathogen |>
  mutate(
    month = month(collectDate),
    year = year(collectDate)
  ) |>
  group_by(siteID, year, month) |>
  summarize(
    positiveCount = n()
  )

ggplot(
  data = daily_positive_pathogen,
  aes(
    x = month, 
    y = positiveCount,
    color = siteID) 
  ) +
  geom_line(size = 0.5) +
theme_bw()


```
