---
title: "Data Exploration Graphs"
date: "2026-03-03"
output: html_document
---

## Load in Packages and Data

Load and install packages that are needed for creating graphs and visualizing data.

```{r}

#install.packages(dplyr)
#install.packages(ggplot2)
#install.packages(here)

library(tidyverse)
library(here)

```

# Temperature Data

Graphs that explore the daily and monthly temperature data sets.

## Data Type Manipulation

Import processed daily temperature information and ensure all dates are date types.

```{r}

daily_temperature_data <- read_csv(here("data","processed","daily_temp_data.csv"))

monthly_temperature_data <- read_csv(here("data", "processed", "monthly_temp_data.csv"))

```

Create year-month in monthly temperature data
```{r}

monthly_temperature_data <- monthly_temperature_data %>% 
  mutate(month_date = make_date(year, month))

```


## Exploratory Graphs

### Daily Mean Temperatures

Line graph visualizing how mean daily temperature changed throughout all dates.

```{r}

ggplot(
  data = daily_temperature_data,
  aes(
    x = startDateTime, 
    y = mean_temp,
    color = siteID) 
  ) +
  #facet_wrap(~siteID) +
  geom_line(size = 0.2) +
theme_bw()

```

### Monthly Mean Temperature

```{r}
#TODO: make line graph similar to graph above but using monthly temperatures

```

### Monthly Temperature Range

```{r}

# Summarize and group monthly temperature data

monthly_temp_range <- monthly_temperature_data |>
  group_by(siteID, month) |>
  summarize(
    mean = mean(mean_temp, na.rm = TRUE),
    max = mean(max_temp, na.rm = TRUE),
    min = mean(min_temp, na.rm = TRUE)
  )

ggplot(
  monthly_temp_range,
  aes(
    x = month,
    y = mean
  )) +
geom_pointrange(
  aes(
    ymin = min,
    ymax = max
  )
) +
  facet_wrap(~siteID)

```
------------------------------------------------------------------------
### Pathogen Status Visualization

```{r}

pathogen_status_data <- read_csv(here("data","processed","pathogen_status_data.csv"))

pathogen_status_data <- pathogen_status_data |>
  mutate(
    collectDate = as_date(collectDate)
  )

monthly_temperature_data <- read.csv(here("data", "processed", "monthly_temp_data.csv"))

```
