---
title: "Statistical Analyses"
authors: "Claire Morgan, Lindsey Stowell, Madison McDonald, Natalie Meyers"
output: html_document
---

# The Relationship Between Seasonal Temperature and Tick Pathogen Abundance

------------------------------------------------------------------------

## Import Packages

```{r}

#install.packages("tidyverse")
#install.packages("here")
#install.packages("dplyr")
#install.packages("lubridate")

library(tidyverse)
library(here)
library(dplyr)
library(lubridate)

```

------------------------------------------------------------------------

## Load Raw Data

```{r}

pathvtemp <- read_csv(here("data","processed","pathogen_vs_temp_data.csv"))

```

## Visualize Temperature vs. Pathogen Count

```{r create figure}

ggplot(
  data = pathvtemp,
  aes(
    x = mean_temp,
    y = pathogenCount)
  ) +
  geom_point() +
  facet_wrap(~siteID) +
theme_bw()

```

## Perform Regression

```{r, linear regression}

ln_model <- lm(mean_temp ~ pathogenCount,
                 data = pathvtemp)

summary(ln_model)

```

## Test assumptions of linear model and diagnostic plots

```{r}
# residuals vs. fitted plot
plot(ln_model, which = 1)

# QQ residuals plot
plot(ln_model, which = 2)

# Scale-location plot
plot(ln_model, which = 3)

# Shapiro test of linear model residuals
shapiro.test(residuals(ln_model))

```

*** note that the p-value from Shapiro test was significant (p<0.05), suggesting that the residuals of the linear model are not normally distributed. lines for 3 plots are also giving wonky so like maybe data doesn't meet assumptions for linear regression, not sure how strict adherence is to 0 horizontal because they're kinda roughly 0?

## Plot Linear Regression

```{r}

lm_plot <- ggplot(pathvtemp,
                  aes(x = mean_temp,
                      y = pathogenCount)) +
  geom_point(alpha = 0.6) + 
  geom_smooth(method = "lm", se = TRUE, color = "seagreen") +
  theme_classic() +
  labs(
    x = "Mean Daily Temperature (ºC)",
    y = "Pathogen Count",
    title = "Relationship of Daily Temperature and Pathogen Count at HARV Site"
  )

lm_plot

```



